Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2022-03-17 paper/paper.tex
update generic programming

\section{Generic programming tensor contractions}\label{sec:generic}
\purple{On a second read of this
section, I think that the clarity could be improved. My understanding is this is the flow:
We would like to use generic programming, where we write code that works regardless
of input type, and is optimized on all input types without sacrificing efficiency. In
general, we can write code such that a function runs correctly regardless of the input
data type. This is easy to do for dynamically typed languages like Python, which only
verify and enforce the rules for the input data type at runtime (called type-checking).
However, when the input data type is only checked at runtime (as opposed to when the
code is compiled), type-specific optimization cannot be used. High efficiency is only
achieved in statically typed languages, where modern compiling technology allows the
function to be individually optimized for each input data type (then give C++/Julia
examples).}
In previous works relating tensor networks and combinatoric problems~\cite{Kourtis2019, Biamonte2017},
the elements in the tensor networks are limited to standard number types such as floating point numbers and integers.
Owning to the development of modern compiling technology, we no longer need to re our imagination to standard number types.
One of the key concepts that push the technology forward is called generic programming:

@@

\purple{I think the following paragraph could be clarified. When you read the definition
of generic programming, it's not clear what ``in the most general setting'' means. My
understanding is it just means that the program will work regardless of the input data
type - it would be good to say this explicitly. I also think the way it is written takes a while
to understand what dynamically versus statically typed languages are. The definition,
“When these languages “see” a new input type, the compiler can recompile the generic
program for the new type.” comes rather late in the paragraph. Also, while I know what
you mean, concepts like “inlining immutable elements” and “cache miss rate” would be
lost on someone only with a physics background.}\blue{This definition is from the book I cited, I also do not like ``in the most general setting'', maybe we just use our version?}
This definition of generic programming contains two major aspects: a single program works in the most general setting and efficiency.
To understand the first aspect on generality, suppose we want to write a function that raises an element to a power, $f(x, n) := x^n$.
One can easily write a function for standard number types that computes the power of $x$ in $O \left( \log(n) \right)$ steps using the multiply and square trick.
Generic programming does not require $x$ to be a standard number type,
instead it treats $x$ as an element with an associative multiplication operation $\odot$ and a multiplicative identity $\mymathbb{1}$.
In such a way, when the program takes a matrix as an input, it computes the matrix power without extra efforts.
The second aspect is about performance. For dynamically typed languages such as Python,
one can easily write very general codes, but the efficiency is not guaranteed; for example, the speed of computing the matrix multiplication between two numpy arrays with python objects as elements is much slower than statically typed languages such as C++ and Julia~\cite{Bezanson2012}.
C++ uses templates for generic programming while Julia takes advantage of just-in-time compilation and multiple dispatch.
When these languages ``see'' a new input type, the compiler can recompile the generic program for the new type.
A myriad of optimizations can be done during the compilation, such as inlining immutable elements with fixed sizes in an array to decrease the cache miss rate when accessing data.
In Julia, these inlined arrays can even be compiled to GPU devices for faster computation~\cite{Besard2018}.

>>>>

\section{Generic programming tensor contractions}\label{sec:generic}
\purple{On a second read of this
section, I think that the clarity could be improved. My understanding is this is the flow:
We would like to use generic programming, where we write code that works regardless
of input type, and is optimized on all input types without sacrificing efficiency. In
general, we can write code such that a function runs correctly regardless of the input
data type. This is easy to do for dynamically typed languages like Python, which only
verify and enforce the rules for the input data type at runtime (called type-checking).
However, when the input data type is only checked at runtime (as opposed to when the
code is compiled), type-specific optimization cannot be used. High efficiency is only
achieved in statically typed languages, where modern compiling technology allows the
function to be individually optimized for each input data type (then give C++/Julia
examples).}\blue{how about now?}
In previous works relating tensor networks and combinatoric problems~\cite{Kourtis2019, Biamonte2017},
the elements in the tensor networks are limited to standard number types such as floating point numbers and integers.
%Owning to the development of modern compiling technology, we no longer need to restrict our imagination to standard number types.
The truth is, the elements in a tensor does not have to be a regular number type,
 or even a field (since division is not used in tensor contraction).
One can use the same tensor contraction program for different purposes with different tensor element types,
and this idea of using the same program for different purposes is also called the generic programming in computer science:

@@

\purple{I think the following paragraph could be clarified. When you read the definition
of generic programming, it's not clear what ``in the most general setting'' means. My
understanding is it just means that the program will work regardless of the input data
type - it would be good to say this explicitly. }
\blue{This definition is from the book I cited, I also do not like ``in the most general setting'', maybe we just use our version?}
\purple{I also think the way it is written takes a while
to understand what dynamically versus statically typed languages are. The definition,
“When these languages “see” a new input type, the compiler can recompile the generic
program for the new type.” comes rather late in the paragraph.}
\xpurple{Also, while I know what you mean, concepts like “inlining immutable elements” and “cache miss rate” would be
lost on someone only with a physics background.}
This definition of generic programming contains two major aspects: a single program works in the most general setting and efficiency.
To understand the first aspect on generality, suppose we want to write a function that raises an element to a power, $f(x, n) := x^n$.
One can easily write a function for standard number types that computes the power of $x$ in $O \left( \log(n) \right)$ steps using the multiply and square trick.
Generic programming does not require $x$ to be a standard number type,
instead it treats $x$ as an element with an associative multiplication operation $\odot$ and a multiplicative identity $\mymathbb{1}$.
In such a way, when the program takes a matrix as an input, it computes the matrix power without extra efforts.
The second aspect is about performance. For dynamically typed languages such as Python,
one can easily write very general codes, but the efficiency is not guaranteed; for example, the speed of computing the matrix multiplication between two numpy arrays with python objects as elements is much slower than statically typed languages such as C++ and Julia~\cite{Bezanson2012}.
C++ uses templates for generic programming while Julia takes advantage of just-in-time compilation and multiple dispatch.
When these languages ``see'' a new input type, the compiler can recompile the generic program for the new type.
A myriad of optimizations can be done during the compilation, such as optimize the memory layout of immutable elements with fixed sizes in an array to speed up the array indexing.
In Julia, arrays with immutable elements with fixed sizes can even be compiled to GPU devices for faster computation~\cite{Besard2018}.
